\paragraph{Acknowledgements}
The theoretical quantum optics group at the TU-Dresden has provided
exemplary conditions for my first foray into the world of scientific
research. I am very grateful for the friendly environment and all the
interesting discussions, as well as the help I received.

I wish to thank Valentin Link for his helpful suggestions that often
put me on the right track. Konstantin Beyer helped me greatly by
pointing out interesting avenues of research and providing helpful
discussion, often being my first ``emergency contact''. He also
encouraged me to attend a seminar to present a poster which proved to
be a wonderful experience. I want to thank Richard Hartmann who helped
me to grasp the HOPS method and always was available to sort out
issues. Without his experience and programming effort none of this
work would have been possible. Kai M\"uller has also been a great and
patient helper and discussion partner. A better office mate wouldn't
be imaginable.

I am grateful for the guidance I received by Prof. Walter Strunz. He
often saw straight through problems and was often able to point out
important questions.

Kimmo Luoma has agreed to serve as the second supervisor of this
thesis, for which I am grateful.

Further thanks to Richard Hartmann, Konstantin Beyer, Kimmo Luoma and
Walter Strunz are in order for proof reading this thesis and providing
very helpful comments. Richard Hartmann was especially thorough at
this task.

I am also grateful to the \href{https://tu-dresden.de/zih}{Center for
  Information Services and High Performance
  Computing}\footnote{Zentrum für Informationsdienste und
  Hochleistungsrechnen (ZIH)} at the TU Dresden for providing its
facilities for high throughput calculations.
